\chapter*{perception}
\addcontentsline{toc}{chapter}{perception}
\begin{center}
\vspace{2cm}
\begin{flushright}
% \large
% \textit{Self-Organized Criticality}
Loud drones, low frequency soothing sounds

Whispers louder than the loudest screams

A new detail that changed my day

The repetitive, unsettling touch

Tight knots, tight hugs

Invasive gazes that were not supposed to last

The faces, the mirrors, the shadows

Acoustics as the language of every surface

\end{flushright}
\vspace{2cm}
% \vspace*{\fill}
\end{center}
\normalsize

Self-Organized Criticality describes how certain systems naturally evolve toward a critical, highly sensitive state where small changes can lead to large-scale effects. This state of criticality is "self-organized" because the system doesn’t require external tuning to reach this point. It naturally arranges itself into this state through its own dynamics. These models help describe the experience of sensory amplification, where the world can be perceived in vivid detail or with overwhelming intensity. \citep{adami1993}

% Suggestion: Add a discussion on how amplified perceptions can highlight decay, such as noticing the breakdown of systems or patterns over time.

For neurodivergent individuals, perception often unfolds in ways that might differ from conventional understandings. Perception is frequently nonlinear and multisensorial, difficult to articulate but deeply felt. The context and balances between anxiety and relaxation shape the sensitivity to sounds and textures. Patterns and structures inan otherwhise chaotic environment is a comforting experience. Time feels non-linear, with moments stretching or compressing, influencing how art and events are experienced. 

As a neurodivergent artist, I often reject traditional forms and mediums, embracing abstraction, fragmentation, or unconventional materials, as an attempt to convey my own personal experiences. I gravitate toward immersive or interactive art. Installation art, for instance, allow for a direct engagement of multiple senses, mirroring how many neurodivergent individuals process the world, translating complex inner experiences into tangible forms.

Sensory overload or hyperfocus can transform the relationship between body and world, and define new ways to conceptualize subjectivity.

From a physics standpoint, perception can be understood as a signal processing system, where sensory organs act as transducers, converting energy from one form to another. Interacting photons, wavelengths of a restricted spectrum, pressure variations converted into electrical signals, molecules binding and interacting with receptors, mechanical interactions of skin cells, electromagnetic repulsion, preventing matter from collapsing into itself. 

Physics reveals that human perception captures only a small fraction of reality. Neutrinos, dark matter, infrared and ultraviolet light, gravitational waves, and radio signals are imperceptible to us, but part of this reality can be detected with specialized instrument. Can the use of such instruments alter our perception? Is it possible to amplify and alter whe ways we relate to the environment by increasing the reach of our perception? 

% post-phenomenology and Umwelt
Technology expands our Umwelt by allowing us to access phenomena that is otherwise absent from our perceptual sphere. From infrared imaging to quantum computers, from microscopes to large language models, technology act as an extension of our perceptual apparatus, expanding our boundaries into previously inaccessible domains.  

According to the phylosofical approach of post-phenomenology, technology is not neutral. It transforms the nature of perception and experience. Instruments shape what and how we perceive, and influence the object of perception as well as the perceiver.

Our brain is remarkably adaptable and capable of incorporating non-biological sensors into its perceptual framework. This phenomenon is supported by research in neuroscience, cognitive science, and philosophy, particularly in the domains of neuroplasticity, sensory substitution, and embodied cognition. The inclusion of non-biological sensors introduces entirely new dimensions to human perception, effectively expanding the "self".

The extended mind hypothesis, proposed by the philosophers Andy Clark and David Chalmers, argues that tools and technologies can become integral parts of our cognitive processes. Non-biological sensors present in wearable technologies provide continuous streams of data, which the brain learns to process and integrate, suggesting that perception is not confined to the brain and body but extends into the tools we use, fundamentally altering our experience.

If perception relies on external tools, the distinction between "human" and "machine" becomes less clear, leading to a hybridized (cyborg) perception that transcends biology. What counts as "real" if our tools mediate all new experiences? Can the brain adapt to perceive entirely artificial data streams, such as simulations or virtual realities, as seamlessly as it does natural environments?

An overwhelming visual stimulation can sometimes be managed via a calming sound, or a specific type of pressure. Since the world is evolving into larger and larger amounts of information and stimulus, tt's interesting to wonder if the inclusion of new types of non-biological sensors to our perception will provoke further overstimulation of present oportunities of relaxation based on new calming sensations. Perhaps soon, focusing on  calming fluctuations of cosmic microwave background radiation will provide a mental shelter from saturated visual and accoustic inputs in our present environment.      


